---
title: "Model Comparison with Bayes Factors"
author: "Ryan Raaum - Spring 2022"
course: "Introduction to Bayesian Statistics"
output:
  rrmdstyle::html_worksheet: default
  rrmdstyle::pdf_worksheet: default
---

## Setup 

You may need to install some packages.

```{r setup-new-packages}
if (!(require("bayestestR", quietly = TRUE))) install.packages("bayestestR")
if (!(require("see", quietly = TRUE))) install.packages("see")
if (!(require("logspline", quietly = TRUE))) install.packages("logspline")
```

As always, load up the packages that we need.

```{r setup-packages, message=FALSE}
library(qmdata) # my example datasets package
theme_set(theme_minimal()) # not necessary, I just don't like the default ggplot theme
library(brms) # for Bayesian modeling
library(dplyr) # for data wrangling
library(bayestestR) # for bayes factor calculations
```

## Setup example data

You'll use the elephant tusk size dataset that you used in earlier worksheets ([Chiyo et al. 2015](https://doi.org/10.1002/ece3.1769)) and we'll do the same data pre-processing as before: logging shoulder height and tusk length and centering the height predictor.

```{r data-pre-processing}
tusksize <- tusksize %>% 
  mutate(loglength = log(length),
         logheight = log(height)) %>% 
  mutate(logheight = scale(logheight, center = TRUE, scale = FALSE)[,1])
```

## Create models

For the model comparisons, we'll fit some of the same models that were compared previously using LOO. First set up the model formulae.

```{r model-formulae}
# one with no interactions
formula1 <- loglength ~ logheight + sex + time

# one including just the two-way interactions that look like they could be non-zero
formula2 <- loglength ~ logheight + sex + time +
  logheight:time + sex:time

# including all two-way interactions
formula3 <- loglength ~ logheight + sex + time +
  logheight:sex + logheight:time + sex:time
```

Now fit the models (we'll also fit an intercept-only model). We'll leave the default intercept prior - because that is a real prior in `brms`, but we'll set the same `normal(0, 2.5)` prior for the Population-level coefficients (instead of the default improper flat prior).

```{r fit-models}
CHAINS = 4
CORES = parallel::detectCores() - 1
SEED = 1234

fit0 <- brm(loglength ~ 1,
           data = tusksize,
           save_pars = save_pars(all = TRUE), # needed for BF comparisons
           chains = CHAINS, cores = CORES, seed = SEED,
           refresh = 0) # suppress the chain output information

fit1 <- brm(formula1,
           data = tusksize,
           prior = prior(normal(0, 2.5), class = b),
           save_pars = save_pars(all = TRUE), # needed for BF comparisons
           chains = CHAINS, cores = CORES, seed = SEED,
           refresh = 0) # suppress the chain output information

# we can reuse all the settings from fit one and just change the formula
# (data needs to be supplied again because there are new terms in the formula)
# (and it seems that the refresh setting doesn't carry over)
fit2 <- update(fit1, formula2, newdata = tusksize, refresh = 0)
fit3 <- update(fit1, formula3, newdata = tusksize, refresh = 0)
```

Now calculate Bayes Factors relative to the intercept-only model using the `bayesfactor_models` function from the `bayestestR` package. (This will suggest more samples, but we can ignore that for this example).

```{r first-bf-comparison}
bayesfactor_models(fit0, fit1, fit2, fit3)
```

Well, the other models are massively better than the intercept-only model (but we probably already guessed that). Let's compare the models with interactions to the model with only main terms.

```{r second-bf-comparison}
bf2 <- bayesfactor_models(fit1, fit2, fit3)
bf2
```

The model with just the `logheight:time` and `sex:time` interactions is substantially better than the one with all the interactions.

There's a kind-of cutesy plot for the model comparison object.

```{r plot-bf-comparison}
plot(bf2)
```

We can also compare the two best fitting models (comparing the better fitting `fit2` against `fit3` to get BF numbers > 1).

```{r compare-best-fitting}
bf3 <- bayesfactor_models(fit3, fit2)
bf3
```

So, by the Bayes Factor guidelines, 27 is a clear advantage of the simpler model over the one with all of the possible two-way interactions.
