---
title: "Hierarchical Multiple Comparisons"
author: "Ryan Raaum - Spring 2022"
course: "Introduction to Bayesian Statistics"
output:
  rrmdstyle::html_worksheet: default
  rrmdstyle::pdf_worksheet: default
---

## Setup 

You'll need to install some packages and install the github version of `brms`.

```{r setup-new-packages}
if (!require("emmeans", quietly = TRUE)) install.packages("emmeans")
if (!require("bayestestR", quietly = TRUE)) install.packages("bayestestR")
if (!require("tidybayes", quietly = TRUE)) install.packages("tidybayes")
```

As always, load up the packages that we need.

```{r setup-packages, message=FALSE}
library(qmdata) # for the example dataset
library(ggplot2) # plotting
library(patchwork) # for laying out plots
theme_set(theme_minimal()) # not necessary, I just don't like the default ggplot theme
library(brms) # for Bayesian modeling
# library(bayesplot) # for plotting of Bayesian models
library(dplyr) # for data wrangling
library(emmeans) # for simple pairwise comparisons and multiple testing correction
library(bayestestR) # for p_direction
library(tidybayes) # for some posterior manipulations
library(posterior) # for other posterior manipulations
# library(posterior) # for working with posterior distributions
```

## Example Data

The example data are from a set of skeletal indices related to functional outcomes from a study examining sexual selection in carnivores [Morris and Carrier 2016](https://doi.org/10.1111/evo.12904). Here we'll focus on the `skull_shape_index` for the Felidae in the dataset.

```{r subset-and-plot-data}
felids <- carnivora_indices %>% 
  filter(family == "Felidae") %>% 
  tidyr::drop_na(skull_shape_index)

ggplot(felids, aes(species, skull_shape_index)) +
  geom_boxplot()
```

## No muliple correction and family-wise error rate correction

We'll fit a regular linear regression model and identify significantly different comparisons without doing any multiple testing correction as well as family-wise error rate correction.

```{r fit-ols-model}
fit <- lm(skull_shape_index ~ species, data=felids)
```

How many of the 15 pairwise comparisons among the six species included are significant at the simple p < 0.05 level?

```{r no-correction}
emmeans(fit, pairwise ~ species, adjust = "none") %>% 
  purrr::pluck("contrasts") %>% 
  tibble::as_tibble() %>% 
  dplyr::filter(p.value < 0.05) %>% 
  nrow()
```

Next, how about a simple Bonferroni correction? (Other options include `holm`, `tukey`, and `fdr`; I couldn't track down the documentation, but it appears to be the `p.adjust` options (see `?p.adjust`) with the addition of Tukey's HSD).

```{r no-correction}
emmeans(fit, pairwise ~ species, adjust = "bonferroni") %>% 
  purrr::pluck("contrasts") %>% 
  tibble::as_tibble() %>% 
  dplyr::filter(p.value < 0.05) %>% 
  nrow()
```

## Bayesian Hierarchical Pairwise Comparisons

We'll fit a hierarchical model with `species` as a group-level variable using `brms`.

```{r fit-hierarchical-model}
CHAINS = 4
CORES = parallel::detectCores() - 1
SEED = 12345

# there's only an intercept and a group-level sd here, 
# which have real priors by default, so will go with those
hfit <- brm(skull_shape_index ~ (1|species),
            data = felids,
            control = list(adapt_delta = 0.9),
            seed = SEED, chains = CHAINS, cores = CORES)
```

### Interlude

The `bayestestR` `p_direction` function does not currently handle `rvars` from the `posterior` package cleanly. The following adds a dispatch function that does that. You can just run this to make it available to use in the next section.

```{r p-direction-for-rvars}
p_direction.rvar <- function(x, method="direct", null=0, ...) {
  x_draws <- posterior::draws_of(x)
  n <- nrow(x_draws)
  pdir <- pmax(
      matrixStats::colCounts(x_draws > null, value=TRUE)/n, # pd positive
      matrixStats::colCounts(x_draws < null, value=TRUE)/n # pd negative
  )
  attr(pdir, "method") <- method
  
  class(pdir) <- unique(c("p_direction", "see_p_direction", class(pdir)))
  
  pdir
}
```

### Back to the example

Do all the pairwise comparisons for each of the group-level variable levels.

```{r hierarchical-comparisons}
hcompare <- hfit %>% 
  spread_rvars(r_species[species,]) %>% 
  compare_levels(r_species, by=species) %>% 
  mutate(pd = p_direction(r_species))
hcompare
```

How many would be designated as different (probability of direction of >= 97.5%)?

```{r hierarchical-comparison-differences}
hcompare %>% 
  filter(pd >= 0.975) %>% 
  nrow()
```

Back to the situation in the class example where the multiple comparison correction might be too extreme when there actually are a lot of real differences in the data.

What do the differences look like?

```{r plot-differences}
hcompare %>% 
  mutate(different = pd >= 0.975) %>%
  ggplot(aes(y = species, dist = r_species, color = different)) +
  stat_dist_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed") 
```

